# Relational data

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
devtools::install_github("bergant/datamodelr")
library(tidyverse)
library(dplyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(fueleconomy)
library(datamodelr)
```

## Exercises

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

-   `flights` table: `origin` and `dest` variables
```{r}
head(flights %>% select(origin, dest))
```
-   `airports` table: faa (matches the origin and destination), longitude and latitude variables

```{r}
head(airports %>% select(lon, lat, faa))
```

-   join `flights` with `airports`. The first join adds the location of the origin airport (`origin`) and the second join adds the location of destination airport (`dest`).

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

The variable `origin` in `weather` is the same as `faa` in `airports`.

```{r}
head(weather %>% select(origin))
head(airports %>% select(faa) %>% filter(faa == "EWR"))
```

`weather` only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

`origin` would need to change to `location`
`year`, `month`, `day`, `hour`, `origin` in `weather` should match to `year`, `month`, `day`, `arr_time`, `dest` in `flight` 

```{r}
head(weather)
head(flights)
```

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

The table `special_days` would contain a list of day where fewer people travel than usual.
Primary key: year, month, day should match `year`, `month`, and `day` in `flight`


```{r}
special_days <- tribble(
  ~year, ~month, ~day, ~holiday,
  2019, 01, 01, "New Years Day",
  2018, 12, 25, "Christmas Day"
)
```

## Exercises 13.3.1 

Add a surrogate key to flights.

```{r}
flights

flights %>%
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) %>%
  glimpse()

flights %>% 
  count(year, month, day, sched_dep_time, carrier, flight) %>% 
  filter(n > 1)
```

Identify the keys in the following datasets

Lahman::Batting

```{r}
head(Lahman::Batting)
Lahman::Batting %>% 
  count(playerID, yearID, stint) %>% 
  filter(n > 1)
```

babynames::babynames

```{r}
head(babynames::babynames)
babynames::babynames %>% 
  count(year, sex, name) %>% 
  filter(nn > 1)
```

nasaweather::atmos

```{r}
head(nasaweather::atmos)
nasaweather::atmos %>%
      count(lat, long, year, month) %>%
      filter(n > 1)
```

fueleconomy::vehicles

```{r}
head(fueleconomy::vehicles)
fueleconomy::vehicles %>%
      count(id) %>%
      filter(n > 1)
```

ggplot2::diamonds

```{r}
head(ggplot2::diamonds)
ggplot2::diamonds %>%
  distinct() %>%
  nrow()
nrow(ggplot2::diamonds)
```

(You might need to install some packages and read some documentation.)

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

-   `Master`

    -   Primary keys: `playerID`

-   `Batting`

    -   Primary keys: `yearID`, `yearID`, `stint`

    -   Foreign Keys:

        -   `playerID` = `Master$playerID` (many-to-1)

-   `Salaries`

    -   Primary keys: `yearID`, `teamID`, `playerID`

    -   Foreign Keys

        -   `playerID` = `Master$playerID` (many-to-1)

```{r}
dm1 <- dm_from_data_frames(list(Batting = Lahman::Batting,
                                Master = Lahman::Master,
                                Salaries = Lahman::Salaries)) %>%
  dm_set_key("Batting", c("playerID", "yearID", "stint")) %>%
  dm_set_key("Master", "playerID") %>%
  dm_set_key("Salaries", c("yearID", "teamID", "playerID")) %>%
  dm_add_references(
    Batting$playerID == Master$playerID,
    Salaries$playerID == Master$playerID
  )

dm_create_graph(dm1, rankdir = "LR", columnArrows = TRUE)
```

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

## Exercises 13.4.6 

Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

airports %>%
  semi_join(flights, c("faa" = "dest")) %>%
  ggplot(aes(lon, lat)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

Add the location of the origin and destination (i.e. the lat and lon) to flights.

Is there a relationship between the age of a plane and its delays?

What weather conditions make it more likely to see a delay?

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.