# Relational data

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
devtools::install_github("bergant/datamodelr")
library(tidyverse)
library(dplyr)
library(nycflights13)
library(babynames)
library(nasaweather)
library(fueleconomy)
library(datamodelr)
```

## Exercises 13.2.1

Imagine you wanted to draw (approximately) the route each plane flies from its origin to its destination. What variables would you need? What tables would you need to combine?

-   `flights` table: `origin` and `dest` variables
```{r}
head(flights %>% select(origin, dest))
```
-   `airports` table: faa (matches the origin and destination), longitude and latitude variables

```{r}
head(airports %>% select(lon, lat, faa))
```

-   join `flights` with `airports`. The first join adds the location of the origin airport (`origin`) and the second join adds the location of destination airport (`dest`).

I forgot to draw the relationship between weather and airports. What is the relationship and how should it appear in the diagram?

The variable `origin` in `weather` is the same as `faa` in `airports`.

```{r}
head(weather %>% select(origin))
head(airports %>% select(faa) %>% filter(faa == "EWR"))
```

`weather` only contains information for the origin (NYC) airports. If it contained weather records for all airports in the USA, what additional relation would it define with flights?

`origin` would need to change to `location`
`year`, `month`, `day`, `hour`, `origin` in `weather` should match to `year`, `month`, `day`, `arr_time`, `dest` in `flight` 

```{r}
head(weather)
head(flights)
```

We know that some days of the year are “special”, and fewer people than usual fly on them. How might you represent that data as a data frame? What would be the primary keys of that table? How would it connect to the existing tables?

The table `special_days` would contain a list of day where fewer people travel than usual.
Primary key: year, month, day should match `year`, `month`, and `day` in `flight`


```{r}
special_days <- tribble(
  ~year, ~month, ~day, ~holiday,
  2019, 01, 01, "New Years Day",
  2018, 12, 25, "Christmas Day"
)
```

## Exercises 13.3.1 

Add a surrogate key to flights.

```{r}
flights

flights %>%
  arrange(year, month, day, sched_dep_time, carrier, flight) %>%
  mutate(flight_id = row_number()) %>%
  glimpse()

flights %>% 
  count(year, month, day, sched_dep_time, carrier, flight) %>% 
  filter(n > 1)
```

Identify the keys in the following datasets

Lahman::Batting

```{r}
head(Lahman::Batting)
Lahman::Batting %>% 
  count(playerID, yearID, stint) %>% 
  filter(n > 1)
```

babynames::babynames

```{r}
head(babynames::babynames)
babynames::babynames %>% 
  count(year, sex, name) %>% 
  filter(nn > 1)
```

nasaweather::atmos

```{r}
head(nasaweather::atmos)
nasaweather::atmos %>%
      count(lat, long, year, month) %>%
      filter(n > 1)
```

fueleconomy::vehicles

```{r}
head(fueleconomy::vehicles)
fueleconomy::vehicles %>%
      count(id) %>%
      filter(n > 1)
```

ggplot2::diamonds

```{r}
head(ggplot2::diamonds)
ggplot2::diamonds %>%
  distinct() %>%
  nrow()
nrow(ggplot2::diamonds)
```

(You might need to install some packages and read some documentation.)

Draw a diagram illustrating the connections between the Batting, Master, and Salaries tables in the Lahman package. Draw another diagram that shows the relationship between Master, Managers, AwardsManagers.

-   `Master`

    -   Primary keys: `playerID`

-   `Batting`

    -   Primary keys: `yearID`, `yearID`, `stint`

    -   Foreign Keys:

        -   `playerID` = `Master$playerID` (many-to-1)

-   `Salaries`

    -   Primary keys: `yearID`, `teamID`, `playerID`

    -   Foreign Keys

        -   `playerID` = `Master$playerID` (many-to-1)

```{r}
dm1 <- dm_from_data_frames(list(Batting = Lahman::Batting,
                                Master = Lahman::Master,
                                Salaries = Lahman::Salaries)) %>%
  dm_set_key("Batting", c("playerID", "yearID", "stint")) %>%
  dm_set_key("Master", "playerID") %>%
  dm_set_key("Salaries", c("yearID", "teamID", "playerID")) %>%
  dm_add_references(
    Batting$playerID == Master$playerID,
    Salaries$playerID == Master$playerID
  )

dm_create_graph(dm1, rankdir = "LR", columnArrows = TRUE)
```

How would you characterise the relationship between the Batting, Pitching, and Fielding tables?

## Exercises 13.4.6 

Compute the average delay by destination, then join on the airports data frame so you can show the spatial distribution of delays. Here’s an easy way to draw a map of the United States:

```{r}
flights_dest_delay <-flights %>% group_by(dest) %>% summarise(avg_arr_delay=mean(arr_delay, na.rm = TRUE)) %>%
  inner_join(airports, by = c(dest = "faa")) %>% select(dest, lon, lat, avg_arr_delay)

flights_dest_delay %>%
  ggplot(aes(lon, lat, colour = avg_arr_delay)) +
    borders("state") +
    geom_point() +
    coord_quickmap()
```
(Don’t worry if you don’t understand what semi_join() does — you’ll learn about it next.)

You might want to use the size or colour of the points to display the average delay for each airport.

Add the location of the origin and destination (i.e. the lat and lon) to flights.

```{r}
airport_locations <- airports %>%
  select(faa, lat, lon)

flights %>%
    select(year:day, hour, origin, dest) %>%
  left_join(
    airport_locations,
    by = c("origin" = "faa")
  ) %>%
  left_join(
    airport_locations,
    by = c("dest" = "faa")
  )
```

Is there a relationship between the age of a plane and its delays?

```{r}
plane_ages<- flights %>% 
  inner_join(planes, by = c("tailnum")) %>%
  mutate(age = year.x - year.y) %>%
  group_by(age) %>%
  summarise(avg_dep_delay = mean(dep_delay, na.rm = TRUE), avg_arr_delay = mean(arr_delay,  na.rm = TRUE)) %>%
  na.omit()

plane_ages
plane_ages %>%
  ggplot(aes(x = age, y = avg_dep_delay)) +
  geom_point() +
  geom_line()

plane_ages %>%
  ggplot(aes(x = age, y = avg_arr_delay)) +
  geom_point() +
  geom_line()
```

What weather conditions make it more likely to see a delay?

```{r}
flight_weather <-
  flights %>%
  inner_join(weather, by = c("origin" = "origin",
                            "year" = "year",
                            "month" = "month",
                            "day" = "day",
                            "hour" = "hour"))

flight_weather %>%
  group_by(temp) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = temp, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(dewp) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = dewp, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(humid) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = humid, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(wind_speed) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = wind_speed, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(wind_gust) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = wind_gust, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(precip) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = precip, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(pressure) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = pressure, y = delay)) +
    geom_line() + 
  geom_point()

flight_weather %>%
  group_by(visib) %>%
  summarise(delay = mean(dep_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  ggplot(aes(x = visib, y = delay)) +
    geom_line() + 
  geom_point()
```

What happened on June 13 2013? Display the spatial pattern of delays, and then use Google to cross-reference with the weather.

In June 13 2013 there was a large series of storms in the southeastern US (see [June 12-13, 2013 derecho series](https://en.wikipedia.org/wiki/June_12%E2%80%9313,_2013_derecho_series))

```{r}
library(viridis)
flights %>%
  filter(year == 2013, month == 6, day == 13) %>%
  group_by(dest) %>%
  summarise(delay = mean(arr_delay, na.rm = TRUE)) %>%
  na.omit() %>%
  inner_join(airports, by = c("dest" = "faa")) %>%
  ggplot(aes(y = lat, x = lon, size = delay, colour = delay)) +
  borders("state") +
  geom_point() +
  coord_quickmap() +
  scale_colour_viridis()
```

## Exercises 13.5.1 

What does it mean for a flight to have a missing tailnum? What do the tail numbers that don’t have a matching record in planes have in common? (Hint: one variable explains ~90% of the problems.)

```{r}
head(flights)
# Flights with tailnum that are not in the planes table
flights %>%
  anti_join(planes, by = "tailnum") %>%
  count(carrier, sort = TRUE)

# tailnum in flight but not in planes
planes %>% filter(tailnum == "N3ALAA")
```

Filter flights to only show flights with planes that have flown at least 100 flights.


```{r}
planes_gt_100 <- flights %>%
  count(tailnum) %>%
  filter(n > 99)

flights %>%
  semi_join(planes_gt_100, by = "tailnum")
```

Combine fueleconomy::vehicles and fueleconomy::common to find only the records for the most common models.

```{r}
glimpse(fueleconomy::vehicles)
glimpse(fueleconomy::common)
```

```{r}
fueleconomy::vehicles %>%
  semi_join(fueleconomy::common, by = c("make", "model"))
```


Find the 48 hours (over the course of the whole year) that have the worst delays. Cross-reference it with the weather data. Can you see any patterns?

```{r}
flights %>%
  group_by(year, month, day) %>%
  summarise(total_24 = sum(dep_delay, na.rm = TRUE)+ sum(arr_delay, na.rm = TRUE)) %>%
  mutate(total_48 = total_24 + lag(total_24)) %>%
  arrange(desc(total_48))
```
What does anti_join(flights, airports, by = c("dest" = "faa")) tell you? What does anti_join(airports, flights, by = c("faa" = "dest")) tell you?


Return all rows from flights where destination (`dest`)  is not matched in airports.  This only keeps columns from flights.
```{r}
# Flights with dest that are not in the airports table
anti_join(flights, airports, by = c("dest" = "faa"))


# dest (`faa`) in flight but not in airports
airports %>% filter(faa == "BQN")
```

```{r}
# Airports that are not in the flights table for dest
anti_join(airports, flights, by = c("faa" = "dest"))

# airport not in dest (`faa`) in flight 
flights %>% filter(dest == "04G")
```

You might expect that there’s an implicit relationship between plane and airline, because each plane is flown by a single airline. Confirm or reject this hypothesis using the tools you’ve learned above.

```{r}
flights %>% group_by(tailnum, carrier)
```


