# Tibbles

## Introduction

No exercises

### Prerequisites

```{r message=FALSE}
library("tidyverse")
```

## Creating Tibbles

No exercises

## Tibbles vs. data.frame

No exercises

## Subsetting

No exercises

## Interacting with older code

No exercises

## Exercises

### Exercise <span class="exercise-number">10.5.1</span> {.unnumbered .exercise}

<div class="question">
How can you tell if an object is a tibble? (Hint: try printing `mtcars`, which is a regular data frame).
</div>

<div class="answer">

`mtcars` prints **all** the columns and it is a `data.frame`.

```{r}
mtcars
```

After converting `mtcars` to a tibble using `as_tibble()`, it prints only the first ten observations.

```{r}
as_tibble(mtcars)
```

Function `is_tibble()` checks whether a data frame is a tibble or not. `mtcars` data frame is not a tibble.

```{r}
is_tibble(mtcars)
```

`diamonds` and `flights` data are tibbles.

```{r}
is_tibble(ggplot2::diamonds)
is_tibble(nycflights13::flights)
is_tibble(as_tibble(mtcars))
```

Function `class()` returns the class of an object. Tibbles has the classes `c("tbl_df", "tbl", "data.frame")`. Data frames has class `"data.frame"`.

```{r}
class(mtcars)
class(ggplot2::diamonds)
class(nycflights13::flights)
```

<!-- If you are interested in reading more on R's classes, read the chapters on -->
<!-- object oriented programming in [Advanced R](http://adv-r.had.co.nz/S3.html). -->

</div>

### Exercise <span class="exercise-number">10.5.2</span> {.unnumbered .exercise}

<div class="question">
Compare and contrast the following operations on a `data.frame` and equivalent tibble. What is different? Why might the default data frame behaviors cause you frustration?
</div>

<div class="answer">

```{r}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

```{r}
tbl <- as_tibble(df)
tbl$x
tbl[, "xyz"]
tbl[, c("abc", "xyz")]
```

Using `$` a data.frame will partially complete the column. So even though we
wrote `df$x` it returned `df$xyz`. This saves a few keystrokes, but can result
in accidentally using a different variable than you thought you were using.

With data.frames, with `[` the type of object that is returned differs on the
number of columns. If it is one column, it won't return a data.frame, but
instead will return a vector. With more than one column, then it will return a
data.frame. This is fine if you know what you are passing in, but suppose you
did `df[ , vars]` where `vars` was a variable. Then you what that code does
depends on `length(vars)` and you'd have to write code to account for those
situations or risk bugs.

</div>

### Exercise <span class="exercise-number">10.5.3</span> {.unnumbered .exercise}

<div class="question">
If you have the name of a variable stored in an object, e.g. `var <- "mpg"`, how can you extract the reference variable from a tibble?
</div>

<div class="answer">

Use the double bracket, like `df[[var]]`. The dollar sign cannot be use because `df$var` would look for a column named `var`.

```{r}
var <- "mpg"
mtcars[[var]]
```

</div>

### Exercise <span class="exercise-number">10.5.4</span> {.unnumbered .exercise}

<div class="question">

Practice referring to non-syntactic names in the following data frame by:

1.  Extracting the variable called 1.
1.  Plotting a scatterplot of 1 vs 2.
1.  Creating a new column called 3 which is 2 divided by 1.
1.  Renaming the columns to one, two and three.

</div>

<div class="answer">

Create a dataset called `practice` with columns named `1` and `2`.

```{r}
practice <- tibble(
  `1` = 1:10,
  `2` = 1:10
)
```

1.  Extract the variable called 1 run

    ```{r}
    practice[["1"]]
    ```

    or

    ```{r}
    practice$`1`
    ```
    
    or

    ```{r}
    one <- "1"
    practice[[one]]
    ```
    
1.  Plotting a scatterplot of 1 vs 2

    ```{r}
    ggplot(practice, aes(x = `1`, y = `2`)) +
      geom_point()
    ```

1.  Creating a new column called 3 which is 2 divided by 1

    ```{r}
    practice[["3"]] <- practice$`2` / practice$`1`
    ```

    or

    ```{r}
    practice[["3"]] <- practice[["2"]] / practice[["1"]]
    ```

    or

    ```{r}
    one <- "1"
    two <- "2"
    three <- "3"
    practice[[three]] <- practice[[two]] / practice[[one]]
    ```
    
1.  To rename the columns to `one`, `two`, and `three`, run

    ```{r}
    practice <- rename(practice, one = `1`, two = `2`, three = `3`)
    glimpse(practice)
    ```

</div>

### Exercise <span class="exercise-number">10.5.5</span> {.unnumbered .exercise}

<div class="question">
What does `tibble::enframe()` do? When might you use it?
</div>

<div class="answer">

Function [`tibble::enframe()`](https://tibble.tidyverse.org/reference/enframe.html) converts named vectors to a data frame with names and values.

The help page for the `enframe()` method of tibble objects is discussed in `?enframe`.


```{r}
enframe(c(a = 1, b = 2, c = 3))
```

</div>

### Exercise <span class="exercise-number">10.5.6</span> {.unnumbered .exercise}

<div class="question">
What option controls how many additional column names are printed at the footer of a tibble?
</div>

<div class="answer">

Function [`print.tbl_df`](https://tibble.tidyverse.org/reference/formatting.html).  The help page for the `print()` method of tibble objects is discussed in `?print.tbl_df`.
The `n_extra` argument determines the number of extra columns to print abbreviated information for.

</div>
