```{r include=FALSE, cache=FALSE}
set.seed(1014)
options(digits = 3)

knitr::opts_chunk$set(
  comment = "#>",
  collapse = TRUE,
  cache = TRUE,
  out.width = "70%",
  fig.align = 'center',
  fig.width = 6,
  fig.asp = 0.618,  # 1 / phi
  fig.show = "hold"
)

options(dplyr.print_min = 6, dplyr.print_max = 6)
```

# Tibble

```{r, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
library(tidyverse)
```

## Exercises

How can you tell if an object is a tibble? (Hint: try printing `mtcars`,
    which is a regular data frame). 

Compare and contrast the following operations on a `data.frame` and 
    equivalent tibble. What is different? Why might the default data frame
    behaviours cause you frustration?

```{r, eval = FALSE}
df <- data.frame(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
    
df <- tibble(abc = 1, xyz = "a")
df$x
df[, "xyz"]
df[, c("abc", "xyz")]    
```

If you have the name of a variable stored in an object, e.g. `var <- "mpg"`,
    how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"
as_tibble(mtcars)[[var]]

as_tibble(mtcars)[var]
```

Practice referring to non-syntactic names in the following data frame by:

```{r}
df <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
```

    Extracting the variable called `1`.
    
```{r}
df[['1']]
df$`1`
```
    
    Plotting a scatterplot of `1` vs `2`.

```{r}
ggplot(df, aes(x = `1`, y = `2`)) +
  geom_point()
```

    Creating a new column called `3` which is `2` divided by `1`.
    
```{r}
df <- df %>% mutate(`3` = `2` / `1`)
```
        
    Renaming the columns to `one`, `two` and `three`. 
    
    ```{r}
df <- df %>% rename(one = `1`, two = `2`, three = `3`)
    ```

What does `tibble::enframe()` do? When might you use it?

```{r}
# enframe() converts named atomic vectors or lists to two-column data frames.
enframe(1:3)
enframe(c(a = 5, b = 7))
```

What option controls how many additional column names are printed at the footer of a tibble?
    
```{r}
options(tibble.width = Inf)
options(tibble.width = 3)
```
