---
title: "Chapter 4: Workflow: basics"
author: "Marjorie Blanco"
date: "12/24/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## 4.4 Practice

Why does this code not work?


```{r}
my_variable <- 10
my_varıable
#> Error in eval(expr, envir, enclos): object 'my_varıable' not found
```

Look carefully! (This may seem like an exercise in pointlessness, but training your brain to notice even the tiniest difference will pay off when programming.)

Ans: An `l` was used instead of `i`

```{r}
my_variable
```

Tweak each of the following R commands so that they run correctly:

```{r, include=FALSE}
library(tidyverse)

ggplot(dota = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

fliter(mpg, cyl = 8)
filter(diamond, carat > 3)
```

```{r}
ggplot(data = mpg) +  # changed dota to data
  geom_point(mapping = aes(x = displ, y = hwy))  

filter(mpg, cyl == 8)   # changed = to ==
filter(diamonds, carat > 3) # changed diamond to diamonds
```

Press Alt + Shift + K. What happens? How can you get to the same place using the menus?

Ans: Runs all code