---
title: "Tibbles"
author: "Marjorie Blanco"
date: "10/8/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

# Creating tibbles

```{r}
as_tibble(iris)

tibble(
  x = 1:5, 
  y = 1, 
  z = x ^ 2 + y
)

tb <- tibble(
  `:)` = "smile", 
  ` ` = "space",
  `2000` = "number"
)
tb

tribble(
  ~x, ~y, ~z,
  #--|--|----
  "a", 2, 3.6,
  "b", 1, 8.5
)
```

# Tibbles vs. data.frame

tibble vs. a classic data.frame: printing and subsetting

## Printing

```{r}
# don’t accidentally overwhelm your console when you print large data frames
tibble(
  a = lubridate::now() + runif(1e3) * 86400,
  b = lubridate::today() + runif(1e3) * 30,
  c = 1:1e3,
  d = runif(1e3),
  e = sample(letters, 1e3, replace = TRUE)
)

nycflights13::flights %>% 
  print(n = 10, width = Inf)

nycflights13::flights %>% 
  View()
```

## Subsetting

```{r}
df <- tibble(
  x = runif(5),
  y = rnorm(5)
)

# Extract by name
df$x

df[["x"]]

# Extract by position
df[[1]]

df %>% .$x

df %>% .[["x"]]
```

# Interacting with older code

```{r}
class(as.data.frame(tb))


```

### Exercises

How can you tell if an object is a tibble? (Hint: try printing mtcars, which is a regular data frame).

```{r}
print(mtcars)
class(mtcars)

as_tibble(mtcars)
class(as_tibble(mtcars))
```


Compare and contrast the following operations on a data.frame and equivalent tibble. What is different? Why might the default data frame behaviours cause you frustration?

```{r}
df <- data.frame(abc = 1, xyz = "a")
df
df$x
df[, "xyz"]
df[, c("abc", "xyz")]
```

If you have the name of a variable stored in an object, e.g. var <- "mpg", how can you extract the reference variable from a tibble?

```{r}
var <- "mpg"
mtcars_tibble <- as_tibble(mtcars)
mtcars_tibble[, var]
```

Practice referring to non-syntactic names in the following data frame by:

```{r}
df <- data.frame(1, 2)
colnames(df) <- c("1", "2")
df
```

Extracting the variable called 1.

```{r}
df$"1"
df[, "1"]
```

Plotting a scatterplot of 1 vs 2.

```{r}
plot(df$"2", df$"1")
plot(df[,"2"], df[,"1"])
```

Creating a new column called 3 which is 2 divided by 1.

```{r}
df$"3" <- df$"2" / df$"1"
df
```

Renaming the columns to one, two and three.

```{r}
colnames(df) <- c("one", "two", "three")
df
```
```{r}
annoying <- tibble(
  `1` = 1:10,
  `2` = `1` * 2 + rnorm(length(`1`))
)
annoying
```

What does tibble::enframe() do? When might you use it?

```{r}
tibble::enframe(1:3)
tibble::enframe(c(a = 5, b = 7))
```

What option controls how many additional column names are printed at the footer of a tibble?
